---
title: "Climatic variables"
author: "Julie K. Sheard"
date: "14 okt 2019"
output: html_document
---

## Comparison of climatic variables

This script compares density distributions of ten climatic variables  for two ant species (Tetramorium immigrans and Tetramorium caespitum).

Climatic variables:
Mean annual temperature
Temperature seasonality
Temperature of warmest quarter
Temperature of coldest quarter
Annual precipitation
Precipitation seasonality
Precipitation of wettest quarter
Precipitation of driest quarter
Precipitation of warmest quarter
Precipitation of coldest quarter

All climatic variables were downloaded from Worldclim 1.4 at 2.5 arc-minutes (  ??5 km) resolution (Hijmans et al., 2005). 


```{r}
#Set theme
theme_set(theme_classic())

#Required libraries
library(plyr)
library(ggplot2)
library(gridExtra)

#Load data
setwd("H:/Documents/Projects/T immigrans/data")
```


```{r}
#Calculate mean value of each climate variable for each species
mu <- ddply(density, "Source", summarise, grp.mean=mean(Annual.Mean.Temperature))
mu2 <- ddply(density, "Source", summarise, grp.mean=mean(Temperature.Seasonality))
mu3 <- ddply(density, "Source", summarise, grp.mean=mean(Mean.Temperature.of.Warmest.Quarter))
mu4 <- ddply(density, "Source", summarise, grp.mean=mean(Mean.Temperature.of.Coldest.Quarter))
mu5 <- ddply(density, "Source", summarise, grp.mean=mean(Annual.Precipitation))
mu6 <- ddply(density, "Source", summarise, grp.mean=mean(Precipitation.Seasonality..Coefficient.of.Variation.))
mu7 <- ddply(density, "Source", summarise, grp.mean=mean(Precipitation.of.Wettest.Quarter))
mu8 <- ddply(density, "Source", summarise, grp.mean=mean(Precipitation.of.Driest.Quarter))
mu9 <- ddply(density, "Source", summarise, grp.mean=mean(Precipitation.of.Warmest.Quarter))
mu10 <- ddply(density, "Source", summarise, grp.mean=mean(Precipitation.of.Coldest.Quarter))
```

###Plot density plots for each climatic variable
Tetramorium immigrans is blue and Tertramoriumn caespitum is orange

```{r}
#Annual mean temperature
one <- ggplot(density, aes(Annual.Mean.Temperature))
onef <- one + geom_density(aes(color=factor(Source)), alpha=0.8) + 
    labs(x= expression(paste(" Annual Mean Temperature [",degree,"C]")), fill="Species") +
    geom_vline(data=mu, aes(xintercept=grp.mean, color=Source),
           linetype="dashed")  + 
    scale_color_manual(values=c("#fc8d59", "#7493c2")) +
    theme(legend.position = "none")

#Temperature seasonality
two <- ggplot(density, aes(Temperature.Seasonality))
twof <-two + geom_density(aes(color=factor(Source)), alpha=0.8) + 
  labs(x= "Temperature Seasonality", fill="Species") +
  geom_vline(data=mu2, aes(xintercept=grp.mean, color=Source),
             linetype="dashed")  + 
  scale_color_manual(values=c("#fc8d59", "#7493c2")) +
  theme(legend.position = "none")

#Temperature of warmest quarter
three <- ggplot(density, aes(Mean.Temperature.of.Warmest.Quarter))
threef <-three + geom_density(aes(color=factor(Source)), alpha=0.8) + 
  labs(x= expression(paste("Mean Temperature of Warmest Quarter [",degree,"C]")), fill="Species") +
  geom_vline(data=mu3, aes(xintercept=grp.mean, color=Source),
             linetype="dashed")  + 
  scale_color_manual(values=c("#fc8d59", "#7493c2")) +
  theme(legend.position = "none")

#Temperature of coldest quarter
four <- ggplot(density, aes(Mean.Temperature.of.Coldest.Quarter))
fourf <-four + geom_density(aes(color=factor(Source)), alpha=0.8) + 
  labs(x= expression(paste("Mean Temperature of Coldest Quarter [",degree,"C]")), fill="Species") +
  geom_vline(data=mu4, aes(xintercept=grp.mean, color=Source),
             linetype="dashed")  + 
  scale_color_manual(values=c("#fc8d59", "#7493c2")) +
  theme(legend.position = "none")

#Annual precipitation
five <- ggplot(density, aes(Annual.Precipitation))
fivef <-five + geom_density(aes(color=factor(Source)), alpha=0.8) + 
  labs(x="Annual Precipitation [mm]", fill="Species") +
  geom_vline(data=mu5, aes(xintercept=grp.mean, color=Source),
             linetype="dashed")  + 
  scale_color_manual(values=c("#fc8d59", "#7493c2")) +
  theme(legend.position = "none")

#Precipitation seasonality
six <- ggplot(density, aes(Precipitation.Seasonality..Coefficient.of.Variation.))
sixf <-six + geom_density(aes(color=factor(Source)), alpha=0.8) + 
  labs(x="Precipitation Seasonality", fill="Species") +
  geom_vline(data=mu6, aes(xintercept=grp.mean, color=Source),
             linetype="dashed")  + 
  scale_color_manual(values=c("#fc8d59", "#7493c2")) +
  theme(legend.position = "none")

#Precipitation of wettest quarter
seven <- ggplot(density, aes(Precipitation.of.Wettest.Quarter))
sevenf <-seven + geom_density(aes(color=factor(Source)), alpha=0.8) + 
  labs(x="Precipitation of Wettest Quarter [mm]", fill="Species") +
  geom_vline(data=mu7, aes(xintercept=grp.mean, color=Source),
             linetype="dashed")  + 
  scale_color_manual(values=c("#fc8d59", "#7493c2")) +
  theme(legend.position = "none")

#Precipitation of driest quarter
eight <- ggplot(density, aes(Precipitation.of.Driest.Quarter))
eightf <-eight + geom_density(aes(color=factor(Source)), alpha=0.8) + 
  labs(x="Precipitation of Driest Quarter [mm]", fill="Species") +
  geom_vline(data=mu8, aes(xintercept=grp.mean, color=Source),
             linetype="dashed")  + 
  scale_color_manual(values=c("#fc8d59", "#7493c2")) +
  theme(legend.position = "none")

#Precipitation of warmest quarter
nine <- ggplot(density, aes(Precipitation.of.Warmest.Quarter))
ninef <-nine + geom_density(aes(color=factor(Source)), alpha=0.8) + 
  labs(x="Precipitation of Warmest Quarter [mm]", fill="Species") +
  geom_vline(data=mu9, aes(xintercept=grp.mean, color=Source),
             linetype="dashed")  + 
  scale_color_manual(values=c("#fc8d59", "#7493c2")) +
  theme(legend.position = "none")

#Precipitation of coldest quarter
ten <- ggplot(density, aes(Precipitation.of.Coldest.Quarter))
tenf <-ten + geom_density(aes(color=factor(Source)), alpha=0.8) + 
  labs(x="Precipitation of Coldest Quarter [mm]", fill="Species") +
  geom_vline(data=mu10, aes(xintercept=grp.mean, color=Source),
             linetype="dashed")  + 
  scale_color_manual(values=c("#fc8d59", "#7493c2")) +
  theme(legend.position = "none")

#Arrange plots in grid
grid.arrange(onef, twof, threef, fourf, fivef, sixf, sevenf, eightf, ninef, tenf, ncol=2)
```

#Compare means to see if species differ in their climatic distribution 

```{r}
#Annual mean temperature
wilcox.test(Annual.Mean.Temperature ~ Source, data = density,
                   exact = FALSE)

#Temperature seasonality
wilcox.test(Temperature.Seasonality ~ Source, data = density,
            exact = FALSE)

#Temperature of warmest quarter
wilcox.test(Mean.Temperature.of.Warmest.Quarter ~ Source, data = density,
            exact = FALSE)

#Temperature of coldest quarter
wilcox.test(Mean.Temperature.of.Coldest.Quarter ~ Source, data = density,
            exact = FALSE)

#Annual precipitation
wilcox.test(Annual.Precipitation ~ Source, data = density,
            exact = FALSE)

#Precipitation seasonality
wilcox.test(Precipitation.Seasonality..Coefficient.of.Variation. ~ Source, data = density,
            exact = FALSE)

#Precipitation of wettest quarter
wilcox.test(Precipitation.of.Wettest.Quarter ~ Source, data = density,
            exact = FALSE)

#Precipitation of driest quarter
wilcox.test(Precipitation.of.Driest.Quarter ~ Source, data = density,
            exact = FALSE)

#Precipitation of warmest quarter
wilcox.test(Precipitation.of.Warmest.Quarter ~ Source, data = density,
            exact = FALSE)

#Precipitation of coldest quarter
wilcox.test(Precipitation.of.Coldest.Quarter ~ Source, data = density,
            exact = FALSE)
```
