---
title: "Ecospat analysis"
author: "Julie Sheard"
date: "11/2/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description

This is the script used to run the niche analyses using the package ecospat

### Load packages

```{r}
library(raster)
library(ade4)
library(ecospat)
library(rgdal)
```

### Load species data 
```{r}
setwd("H:/Documents/Projects/T immigrans/data")

spp <- read.csv("both_wbioclim_clean.csv",h=T,sep=";")
head(spp)
xy.sp1<-subset(spp,species=="sp1")[3:4] #Tetramorium_caespitum
xy.sp2<-subset(spp,species=="sp2")[3:4] #Tetramorium_immigrans
```

### Load and format environmental data, 2.5 minutes

Climatic variables were downloaded from Worldclim

```{r}
setwd("H:/Documents/Projects/T immigrans/data/bio2_2.5_asc")
Bio01 <- raster("__Bio01.asc")
Bio04 <- raster("__Bio04.asc")
Bio12 <- raster("__Bio12.asc")
Bio15 <- raster("__Bio15.asc")
Bio18 <- raster("__Bio18.asc")

env = stack(Bio01, Bio04, Bio12, Bio15, Bio18)
names(env) <- c("bio01", "bio04", "bio12", "bio15", "bio18")
```

### Extract environment for the two species
sp1 = Tetramorium caespitum
sp2 = Tetramorium immigrans

```{r}
env.sp1<-subset(spp,species=="sp1")[6:10]
env.sp2<-subset(spp,species=="sp2")[6:10]

env.sp1<-raster::extract(env,xy.sp1)
env.sp2<-raster::extract(env,xy.sp2)

#Check that headers match between the three environments
head(env.sp1)
head(env.sp2)
head(env)
```

```{r}
#Exclude points with no data
env.sp1.1 <- na.exclude(env.sp1)
env.sp2.2 <- na.exclude(env.sp2)

env.bkg<-na.exclude(values(env))
```


### Ready PCA-environment

```{r}
pca.cal <- dudi.pca(env.bkg, center = TRUE, scale = TRUE, scannf = FALSE, nf = 2)

# predict the scores on the axes
scores.bkg <- na.exclude(pca.cal$li)  #scores for background climate
scores.sp1 <- na.exclude(suprow(pca.cal,env.sp1)$lisup) #scores for sp1
scores.sp2 <- na.exclude(suprow(pca.cal,env.sp2)$lisup) #scores for sp2

# calculation of occurence density (niche z)
z1 <- ecospat.grid.clim.dyn(scores.bkg, scores.bkg, scores.sp1,R=100)
z2 <- ecospat.grid.clim.dyn(scores.bkg, scores.bkg, scores.sp2,R=100)
```

### Carry out niche analysis

```{r}
#Calculate niche overlap metrics D and I
ecospat.niche.overlap(z1,z2 ,cor=T)

#Run niche similarity test
sim.test <- ecospat.niche.similarity.test(z1, z2,
                                          rep=1000, alternative = "greater",
                                          rand.type=2)

#Run niche equivalency test
eq.test <- ecospat.niche.equivalency.test (z1, z2, rep=1000, alternative = "greater", ncores = 1)

dev.off()
par(cex=2.5)
```

### Visualise results

```{r}
#Plot equivalency test
ecospat.plot.overlap.test(eq.test, "D", "Equivalency")
#Plot similarity test
ecospat.plot.overlap.test(sim.test, "D", "Similarity")
```

### Visualise climatic niche overlap in environmental space

```{r}
dev.off()

par(cex=2)
windows()
ecospat.plot.niche.dyn(z1, z2, quant=0.25, interest=2,
                       title= "", name.axis1="PC1",
                       name.axis2="PC2", colZ1="black", colZ2="black",
                       colz2="#7493c2", colz1="#fc8d59")

ecospat.shift.centroids(scores.sp1, scores.sp2, scores.bkg, scores.bkg,col="black")
```

### Conduct PCA for both species combined

```{r}
names(env.sp1.1) <- c("Mean ann. temp.", "Temp. seasonality", "Ann. precip.", "Precip. seasonality", "Precip. warm. quart.")
names(env.sp2.2) <- c("Mean ann. temp.", "Temp. seasonality", "Ann. precip.", "Precip. seasonality", "Precip. warm. quart.")
pca.env <- dudi.pca(rbind(env.sp1.1,env.sp2.2)[,1:5],scannf=F,nf=2)
dev.off()

#Plot variable contributions to PC1 and PC2

windows()
ecospat.plot.contrib(contrib=pca.env$co, eigen=pca.env$eig)
scores <- pca.env$li
```

### Plot species occurrence in geographic space

```{r}
windows()
par(mfrow=c(1,2))

#Tetramorium caespitum
geoz1<-ecospat.niche.zProjGeo(z1,env)
windows()
plot(geoz1,main="Tetramorium caespitum")
points(xy.sp1, pch=20, cex=0.2)

writeRaster(geoz1, filename="TetCae_geo_09022020_new.asc")

#Tetramorium immigrans
geoz2<-ecospat.niche.zProjGeo(z2,env)
plot(geoz2,main="Tetramorium immigrans")
points(xy.sp2, pch=20, cex=0.2)

writeRaster(geoz2, filename="TetImi_geo_09022020_new.asc")

#Clamp map to only show areas with at least 60 % suitability

Tetcae_clamp<- clamp(geoz1,lower=0.6, useValue=TRUE)
Tetimi_clamp<- clamp(geoz2,lower=0.6, useValue=TRUE)
windows()
plot(Tetcae_clamp)
points(xy.sp1, pch=20, cex=1)
windows()
plot(Tetimi_clamp)
points(xy.sp2, pch=20, cex=1)
```